---
title: "p8105_hw1_lk2761"
author: "Lorraine Kwok" 
date: 2019-09-16
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1

This code chunk is used to create a data frame that generates a sample from a normal distribution. 

```{r create data frame for problem 1}

sample_df = tibble(
  sample = rnorm(8),
  gr_th_0 = sample > 0, 
  vec_char_8 = c("This", "generates", "a", "sample", "of", "eight", "random", "numbers"),
  vec_factor_8 = factor(c("odd", "even", "prime", "even", "prime", "even", "odd", "even"))
)

mean(pull(sample_df, sample))

mean(pull(sample_df, gr_th_0)) 

mean(pull(sample_df, vec_char_8))

mean(pull(sample_df, vec_factor_8))

```

This code chunk uses *as.numeric* to convert the logical, character and factor variables from the sample_df data frame to numeric variables. 

```{r applying the as.numeric function to the logical, character and factor variables}

as.numeric(pull(sample_df, gr_th_0))
as.numeric(pull(sample_df, vec_char_8))
as.numeric(pull(sample_df, vec_factor_8))

```

This code chunk is used to: 

  * Convert the logical vector to a numeric vector
  
  * Convert the logical vector to a factor vector
  
  * Convert the logical vector to a factor vector and then to a numeric vector

After each conversion, the random sample will be multipled by that result. 

```{r converting logical vector to other variable types}

gr_th_0_numeric = as.numeric(pull(sample_df, gr_th_0))
gr_th_0_numeric*pull(sample_df, sample)


```

